r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2021-02-24 r-package/paper2020/paper2020.Rmd
add paragraph about 1-1 vs 1-many algorithms

While there are a growing number of open-source routing packages [@opentripplanneropentripplanner; @lovelace2019stplanr; @padgham2019dodgr], most options available do not efficiently process large public transport networks and/or are not user-friendly. To fill this gap, this paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package. R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications, such as accessibility analysis. The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R using seamless parallel computing. This tool can be used to address a variety of questions that require the efficient calculation of travel matrices or the examination of multimodal transport routes.

>>>>

While there are several open-source routing packages available for R, they either do not support public transport networks [@padgham2019dodgr], or focus on providing point-to-point routes rather than origin-destination travel time matrices [@morgan2019opentripplanner; @lovelace2019stplanr]. Most routing algorithms find paths to all points in the network while finding a single route. Storing these paths rather than computing them for one origin-destination pair at a time is orders of magnitude more efficient. To our knowledge, no R package exists that supports these efficient many-to-many queries for public transport networks.

To fill this gap, this paper presents [r5r](https://ipeagit.github.io/r5r/), a new open-source R package for routing on multimodal transport networks based on the [Rapid Realistic Routing on Real-world and Reimagined networks (R<sup>5</sup>)](https://github.com/conveyal/r5) package. R<sup>5</sup> is a powerful next-generation routing engine written in Java and developed at Conveyal [@conway2017evidencebased; @conway2018accounting] to provide an efficient backend for analytic applications, such as accessibility analysis. The r5r package provides a simple and friendly interface to run R<sup>5</sup> locally from within R using seamless parallel computing. This tool can be used to address a variety of questions that require the efficient calculation of travel matrices or the examination of multimodal transport routes.
